---
title: "ADA Final Project"
author: "Lat'Anna Davis and Stephanie Kern-Allely"
date: "April 24, 2019"
output:
  html_document:
    code_folding: "hide"
---

## Add libraries

```{r libraries}
library(knitr) #for creating nicer tables
library(tidyverse) # for various packages
library(dplyr) # FOR PIPING
library(readr) #for read txt file
library(nnet) #Multinomial logistic regression
library(MASS) #Ordinal logistic regression
library(lmtest) #Likelihood ratio test
```

## Add full data set and create short dataset with only variables needed for variable creation
```{r data_loading}

#LOAD FULL DATA
load("34891-0001-Data.rda")

#SELECT VARIABLES FOR HONESTY, TRUST AND RISK OF TERRORISM
data_short<- subset(da34891.0001,select = c(1:2,124:137,
                                       167,170,254:264,
                                       277:278,283:284,
                                       289:290,301:303,
                                       341,343,359:360,
                                       399))  

#REMOVE CREATION DATA AND VARIABLES
rm(da34891.0001)
data_short<-data_short[ -c(3:16)]

#ATTACH THE DATASET
attach(data_short)

```

## Data management to create all, local, state, and federal variables and create analysis data set

```{r data_management}

#CREATE VARIABLES FOR ALL, FED, STATE, LOCAL HONESTY WITH PUBLIC AND CONVERT TO ROUND # FACTORS
all_vars<- sapply(data_short[5:15], as.numeric)
data_short$all_honest<-rowMeans(all_vars,na.rm=TRUE) #average score for all agencies
data_short$all_honest_f<-factor(round(data_short$all_honest), 
                                levels=c(1,2,3,4,5),
                                labels=c("1","2","3","4","5")) #make into round factor variable

local_vars<- sapply(data_short[8:11], as.numeric)
data_short$local_honest<-rowMeans(local_vars,na.rm=TRUE) #average score for all local agencies
data_short$local_honest_f<-factor(round(data_short$local_honest), 
                                  levels=c(1,2,3,4,5),
                                  labels=c("1","2","3","4","5")) #make into round factor variable

state_vars<- sapply(data_short[5:7], as.numeric)
data_short$state_honest<-rowMeans(state_vars,na.rm=TRUE) #average score for all state agencies
data_short$state_honest_f<-factor(round(data_short$state_honest), 
                                  levels=c(1,2,3,4,5),
                                  labels=c("1","2","3","4","5")) #make into round factor variable

fed_vars<- sapply(data_short[12:15], as.numeric)
data_short$fed_honest<-rowMeans(fed_vars,na.rm=TRUE) #average score for all federal agencies
data_short$fed_honest_f<-factor(round(data_short$fed_honest), 
                                levels=c(1,2,3,4,5),
                                labels=c("1","2","3","4","5")) #make into round factor variable

#REMOVE VARIABLES USED FOR CREATION 
rm(all_vars,state_vars,local_vars,fed_vars)
data_short<-data_short[ -c(5:15,30,32,34,36)]

#CREATE AN OVERALL PROTECTION SCORE FROM THE FEDERAL, STATE, LOCAL VARIABLES
# AND CONVERT INTO ROUND FACTOR VARIABLE
all_protect<- sapply(data_short[11:13], as.numeric)
data_short$all_protect<-rowMeans(all_protect,na.rm=TRUE) #average score for all agencies
data_short$all_protect_f<-factor(round(data_short$all_protect), 
                                 levels=c(1,2,3,4,5),
                                 labels=c("1","2","3","4","5")) #make into round factor variable

#REMOVE VECTOR
rm(all_protect)

#CONVERT LIFETIME PERCEIVED RISK OF TERRORISM VARIABLE TO NUMERIC
#data_short$Q7B_2<-as.numeric(Q7B_2)

#CREATE BINARY VARIABLES FOR THINKING AT RISK OF TERRORISM, RENAME VARIABLES FOR LOCAL,
#FED, STATE PROTECTION, WHETHER GOT INFO FROM OFFICIAL SOURCES, EDUCATION, RACE
data_short <- data_short %>%
  mutate( terr_life=Q7B_2,
          #case_when(Q7B_2 == 4 | Q7B_2 == 5 ~ 1, Q7B_2 == 1 | Q7B_2 == 2 | Q7B_2 == 3 ~ 0),
          local_protect=Q9A_2,
          state_protect=Q9A_3,
          fed_protect=Q9A_4,
          believe=Q4E,
          education=Q16,
          race=Q18
          )
#convert NA level to NAs in believe variable
levels(data_short$believe)[levels(data_short$believe)=='(6) N/A'] <- NA
levels(data_short$believe)

#DROP VARIABLES USED TO CALCULATE RISK AND THAT WERE CONVERTED TO NEW NAME (COULD'VE RENAMED 
# IN THE FIRST PLACE BUT DID IT THIS WAY FIRST AND NOT WORTH THE EFFORT TO UNDO)
data_short<-data_short[ -c(2:15,17:18)]

#OLD CODE FOR MAKING FACTORS
#factor_vars <- data_short[c(6:9,13:15,20)]
#data_short[,factor_vars] <-factor(round(data_short[,factor_vars]), 
                               #  levels=c(1,2,3,4,5),
                                # labels=c("1","2","3","4","5"))
#data_short[,factor_vars] <- lapply(data_short[,factor_vars] , factor)

#CATEGORIZE DV AND IVS INTO BINARY VARIABLES
data_short <- data_short %>%
  mutate(all_protect_bin=case_when(all_protect_f == 4 | all_protect_f == 5 ~ 1,
                                   all_protect_f == 1 | all_protect_f == 2 | all_protect_f ==3 ~ 0),
         all_honest_bin=case_when(all_honest_f == 4 | all_honest_f == 5 ~ 1,
                                  all_honest_f == 1 | all_honest_f == 2 | all_protect_f ==3 ~ 0),
         local_protect_bin=case_when(local_protect == "(4) 4" | local_protect == "(5) Extremely sure" ~ 1,
                                     local_protect == "(1) Not at all sure" | local_protect == "(2) 2" | 
                                     local_protect == "(3) 3" ~ 0),
         local_honest_bin=case_when(local_honest_f == 4 | local_honest_f == 5 ~ 1,
                                    local_honest_f == 1 | local_honest_f == 2 | local_honest_f ==3 ~ 0),
         state_protect_bin=case_when(state_protect == "(4) 4" | state_protect == "(5) Extremely sure" ~ 1,
                                     state_protect == "(1) Not at all sure" | state_protect == "(2) 2" | 
                                     state_protect == "(3) 3" ~ 0),
         state_honest_bin=case_when(state_honest_f == 4 | state_honest_f == 5 ~ 1,
                                     state_honest_f == 1 | state_honest_f == 2 | state_honest_f ==3 ~ 0),
         fed_protect_bin=case_when(fed_protect == "(4) 4" | fed_protect == "(5) Extremely sure" ~ 1,
                                   fed_protect == "(1) Not at all sure" | fed_protect == "(2) 2" | 
                                   fed_protect == "(3) 3" ~ 0),
         fed_honest_bin=case_when(fed_honest_f == 4 | fed_honest_f == 5 ~ 1,
                                  fed_honest_f == 1 | fed_honest_f == 2 | fed_honest_f ==3 ~ 0)
  )


#MAKE ANALYSIS DATASET WITH ONLY VARIABLES USED IN ANALYSIS AND REMOVE CASES WITH ANY MISSING
data_analysis<-data_short[c(1:14,16:23)]
data_analysis<-na.omit(data_analysis)

attach(data_analysis)

```

## CHARACTERISTICS OF POPULATION

```{r} 

######
#DEMOGRAPHICS
######

#data_analysis %>% 
 # group_by(all_protect_bin) %>% 
  #summarise(value = median(COMB_AGE), percent = )

#install.packages("table1")
library(table1)

table1(~ factor(education) + COMB_AGE + factor(believe) + factor(terr_life) | all_protect_bin, data=data_analysis)


```

## Perform logistic regression assumptions and overall regression

```{r}


######
#LOGISTIC REGRESSION MODELS
######

#COVARIATE DETERMINATION - DAG

#ASSUMPTIONS

#LINEARITY ASSUMPTION FOR CONTINUOUS PREDICTORS
#AGE
summary(data_analysis$COMB_AGE) #TAKE A LOOK AT THE DISTRIBUTION
shapiro.test(data_analysis$COMB_AGE) #CHECK FOR NORMALITY - RESULT: NOT NORMALLY DISTRIBUTED
log_age <- data_analysis$COMB_AGE * log(data_analysis$COMB_AGE) #MAKE LOGIT TRANSFORMATION
boxTidwellage <- glm(all_protect_bin ~ COMB_AGE + log_age, data=data_analysis, family="binomial") 

#Box Tidwell coefficient is not significant so age does not violate the linearity assumption
#Thus, age is appropriate to run in the logistic model as a continuous variable
summary(boxTidwellage) #RUN BOX TIDWELL TEST FOR AGE ON ALL PROTECT OUTCOME


#INFLUENTIAL VALUES
#logistic model with age as a predictor FOR COOK'S
ageCD <- glm(all_protect_bin ~ COMB_AGE, data=data_analysis, family="binomial")

#There are no values with Cook's Ds above 0.015 but nothing concerning pops out here
#influence plot - Cook's D plot-identifies observation number in parent dataset
plot(ageCD, which=4, id.n=5, col="red") 


#MODELS

#ALL MODEL
#unadjusted
all_model <- glm(all_protect_bin ~ all_honest_bin, data=data_analysis, family = binomial)
summary(all_model)

exp(cbind("Odds ratio" = coef(all_model), confint.default(all_model, level = 0.95)))

#adjusted
all_model_adj <- glm(all_protect_bin ~ all_honest_bin + COMB_AGE + education 
                     + believe + terr_life, 
                     data=data_analysis, family = binomial)
summary(all_model_adj)

OR_all_adj<- exp(cbind("Odds ratio" = coef(all_model_adj), 
                       confint.default(all_model_adj, level = 0.95)))

OR_all_adj[1:2,1:3]

#LogLikelihood - adjusted model is significantly better than the unadjusted model
lrtest(all_model, all_model_adj)

#AFTER ADJUSTING FOR AGE, EDUCATION, BELIEF OF INFORMATION RECEIVED AND LIFETIME PERCEIVED
#RISK OF TERRORISM, THE ODDS OF TRUSTING ALL LEVELS OF GOVERNMENT CAN PROTECT YOU ARE
#23.2 TIMES HIGHER AMONG THOSE WHO BELIEVE THAT ALL LEVELS OF GOVERNMENT ARE HONEST WITH THE PUBLIC
#ABOUT TERRORISM THAN THOSE WHO DO NOT. 

```

## PERFORM logistic regressions for local, state, and federal levels.

```{r}

#LOCAL MODEL
#unadjusted
local_model <- glm(local_protect_bin ~ local_honest_bin, data=data_analysis, family = binomial)
summary(local_model)

OR_local<-exp(cbind("Odds ratio" = coef(local_model), confint.default(local_model, level = 0.95)))

#adjusted
local_model_adj <- glm(local_protect_bin ~ local_honest_bin + COMB_AGE + education 
                     + believe + terr_life, 
                     data=data_analysis, family = binomial)
summary(local_model_adj)

OR_local_adj<-exp(cbind("Odds ratio" = coef(local_model_adj), confint.default(local_model_adj, level = 0.95)))

OR_local_adj[1:2,1:3]

#LogLikelihood - adjusted model is significantly better than the unadjusted model
lrtest(local_model, local_model_adj)

#AFTER ADJUSTING FOR AGE, EDUCATION, BELIEF OF INFORMATION RECEIVED AND LIFETIME PERCEIVED
#RISK OF TERRORISM, THE ODDS OF TRUSTING LOCAL LEVELS OF GOVERNMENT CAN PROTECT YOU ARE
#5.78 TIMES HIGHER AMONG THOSE WHO BELIEVE THAT LOCAL GOVERNMENT IS HONEST WITH THE PUBLIC
#ABOUT TERRORISM THAN THOSE WHO DO NOT. 


#STATE MODEL
#unadjusted
state_model <- glm(state_protect_bin ~ state_honest_bin, data=data_analysis, family = binomial)
summary(state_model)

exp(cbind("Odds ratio" = coef(state_model), confint.default(state_model, level = 0.95)))

#adjusted
state_model_adj <- glm(state_protect_bin ~ state_honest_bin + COMB_AGE + education 
                       + believe + terr_life, 
                       data=data_analysis, family = binomial)
summary(state_model_adj)

OR_state_adj<-exp(cbind("Odds ratio" = coef(state_model_adj), 
                        confint.default(state_model_adj, level = 0.95)))

OR_state_adj[1:2,1:3]

#LogLikelihood - adjusted model is significantly better than the unadjusted model
lrtest(state_model, state_model_adj)

#AFTER ADJUSTING FOR AGE, EDUCATION, BELIEF OF INFORMATION RECEIVED AND LIFETIME PERCEIVED
#RISK OF TERRORISM, THE ODDS OF TRUSTING STATE LEVELS OF GOVERNMENT CAN PROTECT YOU ARE
#5.58 TIMES HIGHER AMONG THOSE WHO BELIEVE THAT STATE GOVERNMENT IS HONEST WITH THE PUBLIC
#ABOUT TERRORISM THAN THOSE WHO DO NOT. 


#FEDERAL MODEL
#unadjusted
fed_model <- glm(fed_protect_bin ~ fed_honest_bin, data=data_analysis, family = binomial)
summary(fed_model)

exp(cbind("Odds ratio" = coef(fed_model), confint.default(fed_model, level = 0.95)))

#adjusted
fed_model_adj <- glm(fed_protect_bin ~ fed_honest_bin + COMB_AGE + education 
                       + believe + terr_life, 
                     data=data_analysis, family = binomial)
summary(fed_model_adj)

OR_fed_adj<-exp(cbind("Odds ratio" = coef(fed_model_adj), 
                      confint.default(fed_model_adj, level = 0.95)))

OR_fed_adj[1:2,1:3]

#LogLikelihood - adjusted model is significantly better than the unadjusted model
lrtest(fed_model, fed_model_adj)

#AFTER ADJUSTING FOR AGE, EDUCATION, BELIEF OF INFORMATION RECEIVED AND LIFETIME PERCEIVED
#RISK OF TERRORISM, THE ODDS OF TRUSTING FEDERAL LEVELS OF GOVERNMENT CAN PROTECT YOU ARE
#4.56 TIMES HIGHER AMONG THOSE WHO BELIEVE THAT FEDERAL GOVERNMENT IS HONEST WITH THE PUBLIC
#ABOUT TERRORISM THAN THOSE WHO DO NOT. 

```